---
title: "Reproducible research - assignment1"
output: html_document
---

This document is created to report works on assignment 1 for course "Reproducible research"



###Loading packages and data (including records with NA)

```{r}
library(ggplot2)
library(plyr)
library(lattice)
url<-"activity.csv"
data0<-read.csv(url, header = TRUE, sep=",")
head(data0)
```

###Removing NAs 

```{r}
dataValid<-data0[!is.na(data0$steps),]
head(dataValid)
```

###Calculating the total number of steps taken by day
```{r}
StepsByDay<-ddply(dataValid,"date",summarise, SumOfSteps=sum(steps))
hist(StepsByDay$SumOfSteps,20,main="Histogram of steps by day")

```

Basic statistics:
```{r}
summary(StepsByDay$SumOfSteps)
```

###Daily activity pattern


```{r}
StepsByInterval<-ddply(dataValid,"interval",summarise, AvgSteps=mean(steps))
plot(StepsByInterval$interval, StepsByInterval$AvgSteps, type="l")
```

Finding interval with maximum average number of steps:
```{r}
maxInterval<-StepsByInterval$interval[StepsByInterval$AvgSteps==max(StepsByInterval$AvgSteps)]
```

Interval with maximum average number of steps: `r maxInterval`

### Imputing missing values

```{r}
stepsNA<-nrow(data0[is.na(data0$steps),])
dateNA<-nrow(data0[is.na(data0$date),])
intervalNA<-nrow(data0[is.na(data0$interval),])
```

Number of NAs in the data set: `r stepsNA` for variable 'steps', `r dateNA` for variable 'date', `r intervalNA` for variable 'interval

Only variable 'steps' needs any dealing with NAs.
Missing values will be replaced with the average for the same interval.

```{r}
dataNAS<-data0[is.na(data0$steps),]
dataNAS<-join(dataNAS,StepsByInterval)
dataNAS$steps<-dataNAS$AvgSteps
dataCleared<-join(dataNAS,dataValid, type="full")
```


Calculating the total number of steps taken by day - for imputed NAs
```{r}
StepsByDay2<-ddply(dataCleared,"date",summarise, SumOfSteps=sum(steps))
hist(StepsByDay2$SumOfSteps,20,main="Histogram of steps by day")

```

Basic statistics:
```{r}
summary(StepsByDay2$SumOfSteps)
```

Only slight differences in distribution of 'steps' by interval, a bit more concentrated around dominant value

### Differences in activity patterns between weekdays and weekends

```{r}
dataCleared$date<-as.Date(dataCleared$date)
dataCleared$weekday<-weekdays(dataCleared$date, abbreviate = TRUE)
dataCleared$weekday2<-"weekday"
dataCleared$weekday2[dataCleared$weekday=="So"|dataCleared$weekday=="N"]<-"weekend"
StepsByWeekday<-ddply(dataCleared,.(interval,weekday2),summarise, MeanOfSteps=mean(steps))
xyplot(MeanOfSteps~interval|weekday2, data=StepsByWeekday, type="l", layout=c(1,2))
```
